---
title: "algatr testing"
format: html
editor: visual
---

20 Feb 2026
following along with the algatr vignette found at https://thewanglab.github.io/algatr/index.html

and using an example lichen metagenome dataset to replace the lizard one they use 


#installing algatr and all the packages needed 

```{r}

devtools::install_github("TheWangLab/algatr")

library(algatr)

# Option 1: Install all of the packages for algatr

alazygatr_packages()

# Option 2: Install subsets of packages based on what methods you want to use:

## Genetic distance processing:

gen_dist_packages()

## Genetic data processing:

data_processing_packages()

## Environmental and geographic data processing:

envirodata_packages()

## LFMM:

lfmm_packages()

## RDA:

rda_packages()

## MMRR:

mmrr_packages()

## GDM:

gdm_packages()

## TESS:

tess_packages()

## wingen:

wingen_packages()

```

#load example lizard dataset

```{r}

load_algatr_example()

```

#look at CA environmental rasters (obtained from World Clim)

```{r}

plot(CA_env, col = turbo(100), axes = FALSE)

```

#combine rasters so you have some multi environmental raster

env <- scaleRGB(CA_env)

plotRGB(env, r = 1, g = 2, b = 3)

# Add sampling localities on top of this (this has to be done outside a codechunk bc it closes the fig within the codechunk and were trying to add points on top of an open fig)

points(liz_coords, pch = 19)

# now going to try to add our own lichen data! - Punctelia appalachensis because it has a reference genome 

but it seems like we need to follow 3 steps in the first 3 scripts from https://github.com/jallen73/Lichen_popgen/tree/main to get vcf files for use in algatr 




the following was done in the terminal 


# step 1 Dowload p. appalachiensis fastq files 

```{bash}
wget ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR344/050/SRR34436550/SRR34436550_1.fastq.gz
wget ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR344/050/SRR34436550/SRR34436550_2.fastq.gz
wget ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR344/051/SRR34436551/SRR34436551_1.fastq.gz
wget ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR344/051/SRR34436551/SRR34436551_2.fastq.gz
```

# rename them 
```{bash}
mv SRR34436550_1.fastq.gz PUAP_HAW_07c_R1_001.fastq.gz
mv SRR34436550_2.fastq.gz PUAP_HAW_07c_R2_001.fastq.gz
mv SRR34436551_1.fastq.gz PUAP_HAW_07b_R1_001.fastq.gz
mv SRR34436551_2.fastq.gz PUAP_HAW_07b_R2_001.fastq.gz
```


```{bash} 
#ignore this 
# mv ~/Desktop/lichen_test/*.fastq.gz ~/Documents/Projects/lichen_specificity/
cd ~/Documents/Projects/lichen_specificity/lichen_specificity
ls -lh *.fastq.gz
```


# step 2 trimming (second script in the bioinformatics pipeline from jessi) run this in the terminal

```{bash}
source /opt/miniconda3/etc/profile.d/conda.sh
conda activate lichen 
```


```{bash}
ls PUAP_*.fastq.gz | sed -e 's/_R._001.*/_R/' | sort | uniq | while read froot ; do
    trimmomatic PE \
        ${froot}1_001.fastq.gz ${froot}2_001.fastq.gz \
        ${froot}1_paired_trimmed.fastq.gz ${froot}1_unpaired_trimmed.fastq.gz \
        ${froot}2_paired_trimmed.fastq.gz ${froot}2_unpaired_trimmed.fastq.gz \
        -threads 4 -phred33 \
        SLIDINGWINDOW:5:20 LEADING:20 TRAILING:20 MINLEN:36
done
```
# now get the reference genome by running this in the terminal 

```{bash}
datasets download genome accession GCA_964256225.1 --include genome
unzip ncbi_dataset.zip
cp ncbi_dataset/data/GCA_964256225.1/GCA_964256225.1_public_SRR14722169_metabat2_bin.6_genomic.fna PUAP_ref.fasta
```

#index it 
```{bash}
bwa index PUAP_ref.fasta
```

#unzip and align
```{bash}
gunzip -k PUAP_HAW_07b_R1_paired_trimmed.fastq.gz
gunzip -k PUAP_HAW_07b_R2_paired_trimmed.fastq.gz
gunzip -k PUAP_HAW_07c_R1_paired_trimmed.fastq.gz
gunzip -k PUAP_HAW_07c_R2_paired_trimmed.fastq.gz

bwa mem PUAP_ref.fasta PUAP_HAW_07b_R1_paired_trimmed.fastq PUAP_HAW_07b_R2_paired_trimmed.fastq > PUAP_HAW_07b.sam
bwa mem PUAP_ref.fasta PUAP_HAW_07c_R1_paired_trimmed.fastq PUAP_HAW_07c_R2_paired_trimmed.fastq > PUAP_HAW_07c.sam
```

```{bash}
samtools view -b PUAP_HAW_07b.sam | samtools sort --threads 4 > PUAP_HAW_07b.bam
samtools view -b PUAP_HAW_07c.sam | samtools sort --threads 4 > PUAP_HAW_07c.bam
```

```{bash}
samtools addreplacerg -r "ID:07b SM:PUAP_HAW_07b" -o PUAP_HAW_07b.rg.bam PUAP_HAW_07b.bam
samtools addreplacerg -r "ID:07c SM:PUAP_HAW_07c" -o PUAP_HAW_07c.rg.bam PUAP_HAW_07c.bam

samtools index PUAP_HAW_07b.rg.bam
samtools index PUAP_HAW_07c.rg.bam
```




#step 3 variance calling (third script in the bioinformatics pipeline from jessi)

#but nothing is working :((((((((()))))))))

```{bash}
#!/bin/bash
## bwa to align reads to reference genome then freebayes to call variants
## run this in your directory with the trimmed fastq files and a reference genome
echo "reference genome is the first argument"
echo "reference genome fasta should end in .fasta"
echo "the file beginning is the second argument"

bwa index $1

ls $2_*.fastq.gz | sed -e 's/\(.*\)_[R][0-9]\{1,2\}_.*$/\1_R/' | sort | uniq | while read froot ; do
    bwa mem $1 ${froot}1_paired_trimmed.fastq.gz ${froot}2_paired_trimmed.fastq.gz -t 20 |
                samtools view -b |
                samtools sort --threads 20 > ${froot}_$(basename $1 .fasta).bam
done

ls *.bam | sort | uniq | while read i ; do
       bamaddrg -b ${i} > $(basename ${i} .bam).rg.bam
done

ls *rg.bam > rgbamlist.txt

ls *.rg.bam | sort | uniq | while read i ; do
       samtools index -b -@ 20 ${i}
done

freebayes -f $1 --min-alternate-fraction 0.3 --min-coverage 10 -p 1 --bam-list rgbamlist.txt > $(basename $1 .fasta)_maf03_mincov10.vcf

##########################################################################################
#Use the following line if you are working with a reference genome that has more than one contig in it
#echo {1..32} | tr " " "\n" | parallel -j 14 "freebayes -f $1 -r contig_{} --min-alternate-fraction 0.5 --min-coverage 10 -p 1 --bam-list rgbamlist.txt > $(basename $1 .fasta)_maf05_mincov10.cont_{}.vcf"

#merge the snps from all vcf files
#ls *.vcf | uniq | while read i; do grep -v '#' $i >> $(basename $1 .fasta)_maf05_mincov10.all.vcf ; done

#check that the sample lists are in the same order for all vcfs
#cat *cont*.vcf | grep '^#C' | uniq | wc -l

#move the header into a single file, then append it to the beginning of the 'all' vcf
#grep '#' $(basename $1 .fasta)_maf05_mincov10.cont_10.vcf > header.vcf
#cat header.vcf $(basename $1 .fasta)_maf05_mincov10.all.vcf > $(basename $1 .fasta)_maf05_mincov10.all.header.vcf
```


