---
title: "algatr testing"
format: html
editor: visual
---

20 Feb 2026
following along with the algatr vignette found at https://thewanglab.github.io/algatr/index.html

and using an example lichen metagenome dataset to replace the lizard one they use 


#installing algatr and all the packages needed 

```{r}

devtools::install_github("TheWangLab/algatr")

library(algatr)

# Option 1: Install all of the packages for algatr

alazygatr_packages()

# Option 2: Install subsets of packages based on what methods you want to use:

## Genetic distance processing:

gen_dist_packages()

## Genetic data processing:

data_processing_packages()

## Environmental and geographic data processing:

envirodata_packages()

## LFMM:

lfmm_packages()

## RDA:

rda_packages()

## MMRR:

mmrr_packages()

## GDM:

gdm_packages()

## TESS:

tess_packages()

## wingen:

wingen_packages()

```

#load example lizard dataset

```{r}

load_algatr_example()

```

#look at CA environmental rasters (obtained from World Clim)

```{r}

plot(CA_env, col = turbo(100), axes = FALSE)

```

#combine rasters so you have some multi environmental raster

env <- scaleRGB(CA_env)

plotRGB(env, r = 1, g = 2, b = 3)

# Add sampling localities on top of this (this has to be done outside a codechunk bc it closes the fig within the codechunk and were trying to add points on top of an open fig)

points(liz_coords, pch = 19)

# now going to try to add our own lichen data! - Punctelia appalachensis because it has a reference genome 

but it seems like there are 3 steps that go from the fastq data file to the vcf file which is what we need to work with algatr 

following the first 3 scripts from https://github.com/jallen73/Lichen_popgen/tree/main to get vcf files for use in algatr 


the following was done in the terminal 


Dowload p. appalachiensis fastq files 

```{bash}
wget ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR344/050/SRR34436550/SRR34436550_1.fastq.gz
wget ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR344/050/SRR34436550/SRR34436550_2.fastq.gz
wget ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR344/051/SRR34436551/SRR34436551_1.fastq.gz
wget ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR344/051/SRR34436551/SRR34436551_2.fastq.gz
```

rename them 
```{bash}
mv SRR34436550_1.fastq.gz PUAP_HAW_07c_R1_001.fastq.gz
mv SRR34436550_2.fastq.gz PUAP_HAW_07c_R2_001.fastq.gz
mv SRR34436551_1.fastq.gz PUAP_HAW_07b_R1_001.fastq.gz
mv SRR34436551_2.fastq.gz PUAP_HAW_07b_R2_001.fastq.gz
```


```{bash}

# mv ~/Desktop/lichen_test/*.fastq.gz ~/Documents/Projects/lichen_specificity/
cd ~/Documents/Projects/lichen_specificity/lichen_specificity
ls -lh *.fastq.gz
```


# step 2 trimming (second script in the bioinformatics pipeline from jessi) run this in the terminal

```{bash}
source /opt/miniconda3/etc/profile.d/conda.sh
conda activate lichen 

```


```{bash}
ls PUAP_*.fastq.gz | sed -e 's/_R._001.*/_R/' | sort | uniq | while read froot ; do
        trimmomatic PE ${froot}1_001.fastq.gz ${froot}2_001.fastq.gz ${froot}1_paired_trimmed.fastq.gz ${froot}1_unpaired_trimmed.fastq.gz ${froot}2_paired_trimmed.fastq.gz ${froot}2_unpaired_trimmed.fastq.gz -threads 4 -phred33 -trimlog ${froot}.log SLIDINGWINDOW:5:20 LEADING:20 TRAILING:20 MINLEN:36
done
```
# now getting the reference genome by running this in the terminal 

datasets download genome accession GCA_964256225.1 --include genome
unzip ncbi_dataset.zip


#step 3 variance calling (third script in the bioinformatics pipeline from jessi)

bwa mem PUAP_ref.fasta PUAP_HAW_07b_R1_paired_trimmed.fastq PUAP_HAW_07b_R2_paired_trimmed.fastq > test.sam
