---
title: "algatr testing"
format: html
editor: visual
---

#installing algatr and all the packages needed 

```{r}

devtools::install_github("TheWangLab/algatr")

library(algatr)

# Option 1: Install all of the packages for algatr

alazygatr_packages()

# Option 2: Install subsets of packages based on what methods you want to use:

## Genetic distance processing:

gen_dist_packages()

## Genetic data processing:

data_processing_packages()

## Environmental and geographic data processing:

envirodata_packages()

## LFMM:

lfmm_packages()

## RDA:

rda_packages()

## MMRR:

mmrr_packages()

## GDM:

gdm_packages()

## TESS:

tess_packages()

## wingen:

wingen_packages()

```

#load example lizard dataset

```{r}

load_algatr_example()

```

#look at CA environmental rasters (obtained from World Clim)

```{r}

plot(CA_env, col = turbo(100), axes = FALSE)

```

#combine rasters so you have some multi environmental raster

env <- scaleRGB(CA_env)

plotRGB(env, r = 1, g = 2, b = 3)

# Add sampling localities on top of this (this has to be done outside a codechunk bc it closes the fig within the codechunk and were trying to add points on top of an open fig)

points(liz_coords, pch = 19)

# now going to try to add our own lichen data!

```{r}

```

